--- 
title: "Morphomics Report"
author: "Shilin Zhao<br><small>Department of Biostatistics<br>Vanderbilt University School of Medicine</small>"
date: "<small>`r Sys.Date()`</small>"
output:
  rmdformats::robobook:
    highlight: kate
    number_sections: no
    code_folding: hide
    toc_depth: 3
    toc_float:
      collapsed: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  markdown:
    wrap: 72
# output:
#   html_document:
#     toc: yes
#     toc_depth: 2
#     number_sections: true
#     # toc_float: 
#     #   collapsed: true
#     # code_folding: hide
#     # theme: cerulean
#     # keep_md: true
description: "some description ..."
---

```{css,echo = FALSE}
.book .book-body .page-inner {
    max-width: 1200px;
}
``` 

```{r setup,echo=FALSE}
#require(Hmisc)    # provides knitrSet and other functions
#knitrSet(lang='markdown', fig.path='png/', fig.align='left', w=6.5, h=4.5, cache=TRUE)
# If using blogdown: knitrSet(lang='blogdown')


knitr::opts_chunk$set(echo = TRUE)
options(width = 3000)

#Hmisc package html Special characters issue
options(htmlSpecialType='&')
```


<!-- # Functions and packages -->

```{r,message=FALSE,warning=FALSE}
library(knitr)
library(tidyverse)

devtools::load_all("D:/source/SpatialPathomicsToolkit")
```


## Load Data

```{r}

dataDir="d:\\workSync\\HaichunYang\\202306_podocyteMorphomics\\202309_data\\"
setwd(dataDir)

files=list.files(pattern="*.csv")
samples=gsub("_.*","",files)
samples=gsub("6793-","",samples)
celltypes=sapply(strsplit(files,"_"),function(x) x[2])

metaTable=data.frame(Sample=samples,File=files,CellType=celltypes,stringsAsFactors=FALSE)
featureDataAll=ReadMorphomicFeatures(metaTable)

```

## Data Transformation
```{r,message=FALSE}

#temp=DataSelection(featureDataAll,featureType=c("AreaShape","Intensity"))
featureDataAllNormlized=MorphomicFeaturesNormlization(featureDataAll)

```


## Data Describe
```{r,message=FALSE}
library(Hmisc)
```

### Raw
```{r,results='asis',eval=FALSE}
varForTable1=colnames(featureDataAllNormlized[["FeatureData"]])
s=describe(featureDataAllNormlized[["FeatureData"]][,varForTable1])
html(s, exclude1=FALSE,  what=c('%'),digits=3, prmsd=TRUE)
```

### Transformation summary
```{r}
print(table(featureDataAllNormlized[["FeatureDataNormlizedSummary"]]$note))
```

### After data Transformation
```{r,results='asis'}
varForTable1=colnames(featureDataAllNormlized[["FeatureDataNormlized"]])
s=describe(featureDataAllNormlized[["FeatureDataNormlized"]][,varForTable1])
html(s, exclude1=FALSE,  what=c('%'),digits=3, prmsd=TRUE)
```

## Perform PCA and Correlation 
```{r,results='hide'}
featureDataAllNormlized=featureDataPcaAll(featureDataAllNormlized)
```

## PCA: Propotion of Variance Explained
### By Feature Type
```{r,fig.width=9,fig.height=9,message=FALSE}
pList=list()
for (reductionName in c("AreaShape","Granularity"  ,"Intensity","Neighbors", "RadialDistribution" ,"Texture"  )){
  pList[[reductionName]]=plotPcaBarplot(featureDataAllNormlized,reductionName=paste0(reductionName,"_PCA"))+ggtitle(reductionName)
}

#show plots in pList by patchwork package
patchwork::wrap_plots(pList,ncol=2)
#patchwork::wrap_plots(pList,ncol=2,guides='collect')&scale_fill_gradientn(colors = color_palette,limits=c(0,30))


```

### By Cell Type
```{r,message=FALSE}
pList=list()
for (reductionName in c("mesangial","PEC", "podocyte" ,"ALL"  )){
  pList[[reductionName]]=plotPcaBarplot(featureDataAllNormlized,reductionName=paste0(reductionName,"_PCA"))+ggtitle(reductionName)
}
patchwork::wrap_plots(pList,ncol=2)

```


## PCA: Correlation of PCs in different cell
### By Feature Type
```{r,fig.width=9,fig.height=9}
# names(featureDataAllNormlized[["Reductions"]])
# 
# reductionName="Texture"
# temp1=featureDataAllNormlized[["Reductions"]][[reductionName]]$cell.embeddings[,3]
# 
# reductionName="podocyte_Texture"
# temp2=featureDataAllNormlized[["Reductions"]][[reductionName]]$cell.embeddings[,3]

plotPCsByCellType(featureDataAllNormlized,reductionName="AreaShape_PCA")
plotPCsByCellType(featureDataAllNormlized,reductionName="Granularity_PCA")
plotPCsByCellType(featureDataAllNormlized,reductionName="Intensity_PCA")
plotPCsByCellType(featureDataAllNormlized,reductionName="Neighbors_PCA")
plotPCsByCellType(featureDataAllNormlized,reductionName="RadialDistribution_PCA")
plotPCsByCellType(featureDataAllNormlized,reductionName="Texture_PCA")

```

### All Features together
```{r,fig.width=12,fig.height=9}
plotPCsByCellType(featureDataAllNormlized,reductionName="ALL_PCA",reductionNameToComp=c("mesangial_PCA","PEC_PCA" ,"podocyte_PCA"),PCs=1:9)

# plotPCsByCellType(featureDataAllNormlized,reductionName="ALL",reductionNameToComp=c("AreaShape","Granularity" ,"Intensity"),PCs=1:4)

```

## PCA: Loading of PCs in different cell
```{r,fig.width=9,fig.height=9}
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="AreaShape_PCA")
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="Granularity_PCA")
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="Intensity_PCA")
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="Neighbors_PCA")
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="RadialDistribution_PCA")
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="Texture_PCA")

```

### All Features together
```{r,fig.width=9,fig.height=20}
plotPCsLoadingByCellType(featureDataAllNormlized,reductionName="ALL_PCA",reductionNameToComp=c("mesangial_PCA","PEC_PCA" ,"podocyte_PCA"),PCs=1:9)

```


## Feature Correlation Heatmap
```{r,fig.width=9,fig.height=9,message=FALSE}
plotCorHeatmap(featureDataAllNormlized,PCs=1:4)
plotCorHeatmap(featureDataAllNormlized,reductionName="podocyte_PCA",PCs=1:4)
plotCorHeatmap(featureDataAllNormlized,reductionName="Intensity_PCA",PCs=1:4)
plotCorHeatmap(featureDataAllNormlized,reductionName="podocyte_Intensity_PCA",PCs=1:4)

```

## sparsePCA analysis
### Perform sparsePCA based on PCA results
```{r}
#another package
#devtools::install_github("erichson/spca")
library(sparsepca)


temp=featureDataAllNormlized
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=30,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$ALL_SparsePCA=resultSparsePCA


featureType="AreaShape"
temp=DataSelection(featureDataAllNormlized,featureType=featureType)
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=10,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$AreaShape_SparsePCA=resultSparsePCA

featureType="Intensity"
temp=DataSelection(featureDataAllNormlized,featureType=featureType)
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=4,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$Intensity_SparsePCA=resultSparsePCA

featureType="Neighbors"
temp=DataSelection(featureDataAllNormlized,featureType=featureType)
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=4,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$Neighbors_SparsePCA=resultSparsePCA

featureType="RadialDistribution"
temp=DataSelection(featureDataAllNormlized,featureType=featureType)
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=10,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$RadialDistribution_SparsePCA=resultSparsePCA

featureType="Texture"
temp=DataSelection(featureDataAllNormlized,featureType=featureType)
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=4,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$Texture_SparsePCA=resultSparsePCA

featureType="Granularity"
temp=DataSelection(featureDataAllNormlized,featureType=featureType)
resultSparsePCA=featureDataSparsePca(temp,dataName="FeatureDataNormlized",k=4,alpha = 1e-03)
featureDataAllNormlized[["Reductions"]]$Granularity_SparsePCA=resultSparsePCA


# 
# dataForsPCA=temp[["FeatureDataNormlized"]]
# set.seed(123)
# rspca.results <- rspca((dataForsPCA), k=10, verbose=TRUE, max_iter=1000,center=TRUE, scale=TRUE)
# row.names(rspca.results$loadings)=colnames(dataForsPCA)
# head(rspca.results$loadings)
# 
# summary(rspca.results)
# 
# 
# featureType="Intensity"
# temp=DataSelection(featureDataAllNormlized,featureType=featureType)
# dataForsPCA=temp[["FeatureDataNormlized"]]
# 
# rspca.results <- rspca((dataForsPCA), k=5, verbose=TRUE, max_iter=1000,center=TRUE, scale=TRUE,alpha = 1e-03)
# row.names(rspca.results$loadings)=colnames(dataForsPCA)
# colnames(rspca.results$loadings)=paste0("PC",1:ncol(rspca.results$loadings))
# 
# head(rspca.results$loadings)
# #how many features for each PC
# apply(rspca.results$loadings,2,function(x) length(which(x!=0))/length(x))
# #Existing in how many PCs for each feature
# table(apply(rspca.results$loadings,1,function(x) length(which(x!=0))))
# 
# summary(rspca.results)

# temp=rspca.results$loadings
# temp[temp==0]=NA
# ComplexHeatmap::Heatmap(temp,cluster_columns = FALSE,cluster_rows =  FALSE)




```

### sparsePCA loading Vis
```{r,fig.width=9,fig.height=9,warning=FALSE}
for (reduction in c("ALL_SparsePCA","AreaShape_SparsePCA","RadialDistribution_SparsePCA")) {
  plot(plotPCsLoading(featureDataAllNormlized[["Reductions"]][[reduction]]$feature.loadings[,1:9])+ggtitle(reduction))
}

for (reduction in c("Intensity_SparsePCA","Neighbors_SparsePCA","Texture_SparsePCA","Granularity_SparsePCA")) {
  plot(plotPCsLoading(featureDataAllNormlized[["Reductions"]][[reduction]]$feature.loadings[,1:4])+ggtitle(reduction))
}

# plotPCsLoading(featureDataAllNormlized[["Reductions"]]$AreaShape_SparsePCA$feature.loadings[,1:9])
# plotPCsLoading(featureDataAllNormlized[["Reductions"]]$RadialDistribution_SparsePCA$feature.loadings[,1:9])
# plotPCsLoading(featureDataAllNormlized[["Reductions"]]$Intensity_SparsePCA$feature.loadings[,1:4])
# plotPCsLoading(featureDataAllNormlized[["Reductions"]]$Neighbors_SparsePCA$feature.loadings[,1:4])
# plotPCsLoading(featureDataAllNormlized[["Reductions"]]$Texture_SparsePCA$feature.loadings[,1:4])
# plotPCsLoading(featureDataAllNormlized[["Reductions"]]$Granularity_SparsePCA$feature.loadings[,1:4])



```



## Differential detection
### Diff By cell type in all samples
```{r,results='asis'}
testResult=testFeatureDiff(featureDataAllNormlized,dataName="AreaShape_PCA",groupColumn="CellType")
knitr::kable(head(testResult))
plotFeatureByGroup(featureDataAllNormlized,dataName="AreaShape_PCA",featureNames=c("PC1","PC2","PC4","PC5"),groupColumn="CellType")

```

### Diff By Sample in one cell type
```{r,results='asis'}
temp=DataSelection(featureDataAllNormlized,CellType="mesangial")
testResult=testFeatureDiff(temp,dataName="AreaShape_PCA",groupColumn="Sample")
knitr::kable(head(testResult))
plotFeatureByGroup(temp,dataName="AreaShape_PCA",featureNames=c("PC2","PC5"),groupColumn="Sample")

#extract cells with largest PC2 and smallest PC2
cellEmbeddingsTable=temp[["Reductions"]][["AreaShape_PCA"]]$cell.embeddings[row.names(temp[["SampleData"]]),]

temp1=cellEmbeddingsTable[head(order(cellEmbeddingsTable[,"PC2"])),][,1:3]
temp2=cellEmbeddingsTable[tail(order(cellEmbeddingsTable[,"PC2"])),][,1:3]

```


### Diff By Group in each cell type
```{r,results='asis'}
featureDataAllNormlized[["SampleData"]]$Group=ifelse(featureDataAllNormlized[["SampleData"]]$Sample %in% c("AF-7","AF-8"),"Sample78","Sample910")

allCellTypes=unique(featureDataAllNormlized[["SampleData"]]$CellType)
testResultList=list()
for (cellType in allCellTypes) {
  temp=DataSelection(featureDataAllNormlized,CellType=cellType)
  testResult=testFeatureDiff(temp,dataName="AreaShape_PCA",groupColumn="Group")
  #knitr::kable(head(testResult))
  testResultList[[cellType]]=testResult
}

for (cellType in names(testResultList)) {
  print(knitr::kable(head(testResultList[[cellType]]),caption = cellType))
}

temp=DataSelection(featureDataAllNormlized,CellType="mesangial")
plotFeatureByGroup(temp,dataName="AreaShape_PCA",featureNames=c("PC2"),groupColumn="Group")

temp=DataSelection(featureDataAllNormlized,CellType="PEC")
plotFeatureByGroup(temp,dataName="AreaShape_PCA",featureNames=c("PC2","PC6"),groupColumn="Group")

temp=DataSelection(featureDataAllNormlized,CellType="podocyte")
plotFeatureByGroup(temp,dataName="AreaShape_PCA",featureNames=c("PC1","PC2","PC4"),groupColumn="Group")

#PC2 is more related
#extract cells with largest PC2 and smallest PC2
cellEmbeddingsTable=featureDataAllNormlized[["Reductions"]][["AreaShape_PCA"]]$cell.embeddings
temp1=cellEmbeddingsTable[head(order(cellEmbeddingsTable[,"PC2"])),][,1:3]
temp2=cellEmbeddingsTable[tail(order(cellEmbeddingsTable[,"PC2"])),][,1:3]

#dput(row.names(temp1))
#dput(row.names(temp2))
```




